---
title: "Linear Programming"
author: "Dutt Thakkar"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Linear Programming Problem

# Formulating LP problem and writing constraints.
# *Objective function*

# Max: 420(X1+Y1+Z1)+360(X2+Y2+Z2)+300(X3+Y3+Z3);

# *Plant Production Constraint*

# X1+X2+X3 <= 750;
# Y1+Y2+Y3 <= 900;
# Z1+Z2+Z3 <= 450;

# *In-process Storage Constraint*

# 20X1+15X2+12X3 <= 13000;
# 20Y1+15Y2+12Y3 <= 12000;
# 20Z1+15Z2+12Z3 <= 5000;

# *Sales Constraint*

# X1+Y1+Z1 <= 900;
# X2+Y2+Z2 <= 1200;
# X3+Y3+Z3 <= 750;

# *Same Percentage Constraint* 

# (1/750)*(X1+X2+X3)-(1/900)*(Y1+Y2+Y3)=0;
# (1/750)*(X1+X2+X3)-(1/450)*(Z1+Z2+Z3)=0;

# Installing and loading the lpsolve package

```{r}
library(lpSolve)
```

# Coefficients of the objective function

```{r}
objective_c <- c(420, 360, 300, 420, 360, 300, 420, 360, 300)
```

# Setting the inequality constraint matrix

```{r}
const_mat <- matrix(c
                (1, 0, 0, 1, 0, 0, 1, 0, 0,
                 0, 1, 0, 0, 1, 0, 0, 1, 0,
                 0, 0, 1, 0, 0, 1, 0, 0, 1,
                 1, 1, 1, 0, 0, 0, 0, 0, 0,
                 0, 0, 0, 1, 1, 1, 0, 0, 0,
                 0, 0, 0, 0, 0, 0, 1, 1, 1,
                 20, 15, 12, 0, 0, 0, 0, 0, 0,
                 0, 0, 0, 20, 15, 12, 0, 0, 0,
                 0, 0, 0, 0, 0, 0, 20, 15, 12,
                 900, 900, 900, -750, -750, -750, 0, 0, 0,
                 0, 0, 0, 450, 450, 450, -900, -900, -900,
                 450, 450, 450, 0, 0, 0, -750, -750, -750),
               nrow=12,
               byrow=TRUE)
```

# Directions of constraints 

```{r}
const_dir <- c("<=","<=","<=","<=","<=","<=","<=","<=","<=","=","=","=")
```

# Right hand side of constraints 

```{r}
const_rhs <- c(900,1200,750,750,900,450,13000,12000,5000,0,0,0)
```

# Solving LP problem

```{r}
Z = lp("max", objective_c, const_mat, const_dir, const_rhs)
```

# Printing variable values

```{r}
cat("Optimal Solution (X1, X2, X3, Y1, Y2, Y3, Z1, Z2, Z3):\n", Z$solution, "\n")
cat("Optimal Profit (P):\n", Z$objval, "\n")
```
# Optimal Profit = 696000

# Number of large, medium, and small sized products to be produced at plant 1
# X1 = 516.6667
# X2 = 177.7778
# X3 = 0

# Number of large, medium, and small sized products to be produced at plant 2
# Y1 = 0
# Y2 = 666.6667
# Y3 = 166.6667

# Number of large, medium, and small sized products to be produced at plant 3
# Z1 = 0
# Z2 = 0
# Z3 = 416.6667


